### 2015-09-09 bioc192/lect25.tex
finished lecture 25, onto lecture 26

\begin{center}
\ch{Glucose + Pi ->[Hexokinase] Glucose-6-phosphate + H2O}
\end{center}

>>>>

\begin{center}
    \ch{Glucose + Pi ->[Hexokinase] Glucose{-} 6 {-}phosphate + H2O}
\end{center}

The reaction above has a \bioGibbs[exponent=0']{+14}, which means that the reaction is unfavourable.
However, this reaction can occur by coupling this reaction with ATP hydrolysis to make the overall \bioGibbs[exponent=0']{-16}.

\subsection{Oxidising Fuels for Energy}

Both carbohydrates and fats release energy, but we don't want all of the energy to be released/burnt at once.
Therefore, our body requires a stepwise oxidation of carbohydrates and fats to capture the energy.

In your body, there are 2 main kinds of reactions involved in energy metabolism: redox reactions and phosphorylation reactions.

\subsubsection{Redox Reactions}

Redox reactions are where the fuel molecules get oxidised, and the energy from this oxidation is captured by the co-enzymes which gets reduced from this reaction.
Eventually, these reduced co-enzymes are used to make ATP.

\subsubsection{Phosphorylation Reactions}

Phosphorylation reactions are for making/generating ATP by directly or indirectly phosphorylating the ADP.
There are two kinds of phosphorylation reactions:
\begin{description}
    \item[Substrate level phosphorylation] This is where the energy to phosphorylate the ATP comes directly from the reactions that break down the molecules.
        Substrate level phosphorylation occurs in glycolysis and citric acid cycle.
    \item [Oxidative phosphorylation] This is where the energy to phosphorylate the ATP comes indirectly from the breakdown of the molecules, via co-enzymes.
        In the redox reactions, the co-enzymes get reduced, and the energy from these reduced co-enzymes can be fed into oxidative phosphorylation.
\end{description}

\subsection{Reducing Equivalents}

\begin{center}
    \ch{H = H+ + e-}
\end{center}
Biological reactions often involve the transfer of hydrogens (protons + electrons), instead of just electrons in purely chemical reactions.
The co-enzymes NAD and FAD interact with various enzymes to accept/donate H$^+$/e$^-$.
Enzymes that catalyse these reactions are often called dehydrogenase, as it removes hydrogen.

\subsection{Co-enzymes}

\subsubsection{NAD}

NAD (nicotinamide adenine dinucleotide) is made up of niacin (vitamin B3).
NAD accepts hydrogens and electrons in metabolic pathways such as glycolysis, fatty acid oxidation, and citric acid cycle.

\begin{center}
\ch{NAD+ + 2 H+ + 2 e- <=> NADH + H+}\\
\vspace{0.5cm}
\setatomsep{2em}
NAD$^+$ + \chemfig{R-C(-[2]H)(-[6]OH)-R'}
\ch{<=>} NADH + \chemfig{R-C(=[6]O)-R'}
+ H$^+$
\end{center}

As you can see, NAD$^+$ gains a hydrogen atom and two electrons(hence no charge after the reaction).

\subsubsection{FAD}

FAD (flavin adenine dinucleotide) is made from riboflavin (vitamin B2), and it accepts hydrogens in pathways such as fatty acid oxidation and citric acid cycle.
Flavin co-enzymes are tightly bound to the proteins with which they interact, whereas NAD is free in solution and dissociates after the reaction.

\begin{center}
    \ch{ FAD + 2 H+ + 2 e- <=> FADH2 }\\
    \vspace{0.5cm}
    \setatomsep{2em}
    FAD + \chemfig{R-C(-[2]H)(-[6]H)-C(-[2]H)(-[6]H)-R'}
    \ch{<=> FADH2} +
    \chemfig{R-C(-[2]H)=C(-[2]H)-R'}
\end{center}

\subsubsection{Co-enzyme A}

Co-enzyme A (CoA) are made from pantothenic acid (vitamin B5), and acts as a carrier of acyl groups (i.e. a fatty acid carrier).
There are 2 forms: CoASH form (free form) and AcCoA (acyl group attached) form.

\begin{center}
    \setatomsep{2em}
    \ch{H+ + CoASH +} \chemfig{CH_3-C(=[2]O)-\chemabove{O}{\scrm}}
    \ch{<<=> H2O + } \chemfig{CH_3-C(=[2]O)-SCoA}
\end{center}
